IML Term project

### 2022-12-07 main_markdown.rmd

## accuracy
accuracy <- function(pred, y) mean(ifelse(pred >= 0.5, 1, 0) == y)

## perplexity. Force probabilities to [0.995, 0.005]
perplexity <- function(pred, y){
  pred <- ifelse(pred != 0 & pred != 1, pred, ifelse(pred == 0, 0.005, 0.995 ))
  exp(-mean(log(ifelse(y == 1, pred, 1 - pred)))) }

# cross-validate
# split n items into k folds of roughly equal size
kpart <- function(n, k) {
    rep_len(1:k, length.out = n)
}

# Cross-validation predictor (from week 1 problem 2)¨
cv <- function(
    formula, # variables to use
    data, # data
    model = lm, # model to train
    n = nrow(data), # rows in the matrix
    k = min(n, 10), # cross-validation folds
    split = kpart(n, k), # the split of n data items into k folds
    train = function(data) model(formula, data = data),
    pred = function(model, data) predict(model, newdata = data)) {
  yhat <- NULL
  for (i in 1:k) {
    mod <- train(data[split != i, ])
    if (is.null(yhat)) {
      yhat <- pred(mod, data)
    } else {
      yhat[split == i] <- pred(mod, data[split == i, ])
    }
  }
  return(yhat)
}

@@

```{r correlation at different heights, message=FALSE}
# Correlation matrix
cm <- cor(npf[, endsWith(colnames(npf), ".mean")])
# Remove .mean from column and row names.
colnames(cm) <- rownames(cm) <- sapply(colnames(cm), function(s) gsub(".mean", "", s))
# Correlations for each variables (when more than one variable with the same beginning)
kable(cm[startsWith(colnames(cm), "CO2"), startsWith(colnames(cm), "CO2")],
      caption = "Correlation (CO2)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "H2O"), startsWith(colnames(cm), "H2O")],
      caption = "Correlation (H20)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NO)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NOx)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "O3"), startsWith(colnames(cm), "O3")],
      caption = "Correlation (O3)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "RHIRGA"), startsWith(colnames(cm), "RHIRGA")],
      caption = "Correlation (RHIRGA)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "T"), startsWith(colnames(cm), "T")],
      caption = "Correlation (T)") %>%
  kable_styling(latex_options = "HOLD_position")
```

@@

```{r measurementHeights, message=FALSE}
# Selecting which measurements have been done at particular heights
# Columns
cols <- unname(sapply(colnames(npf[endsWith(colnames(npf), ".mean")]), function(s) gsub(".mean", "", s)))
# Table of variables at different heights (only the variables that have been measured at different heights)
kable(data.frame(dm.42 = c(sort(cols[endsWith(cols, "42")])[1:6], NA, sort(cols[endsWith(cols, "42")])[7]),
                                dm.84 = c(NA, sort(cols[endsWith(cols, "84")])[1:5], NA, sort(cols[endsWith(cols, "84")])[6]),
                                dm.168 = c(sort(cols[endsWith(cols, "168")])),
                                dm.336 = c(sort(cols[endsWith(cols, "336")])[1:4], NA, sort(cols[endsWith(cols, "336")])[5], NA, NA),
                                dm.504 = c(sort(cols[endsWith(cols, "504")])[1:6], NA, sort(cols[endsWith(cols, "504")])[7]),
                                dm.672 = c(NA, NA, sort(cols[endsWith(cols, "672")])[2:5], NA, sort(cols[endsWith(cols, "672")])[6])),
      caption = "Measurements at different heights: What measurements have been done at particular heights?") %>%

  kable_styling(latex_options = "HOLD_position")
```

### vif
```{r vif}
library(car)

testmod<-glm(class2~., npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod <- glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std, npf.168,family=binomial(link=logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean, npf.168,family=binomial(link=logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std, npf.168,family=binomial(link=logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std-NOx168.mean-NOx168.std, npf.168,family=binomial(link=logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
```

Runnig vif() over several variables in the 16.8m-restricted dataset, we come up to the following collection of coefficients 
```{r vif coefficients}
coefficients(testmod)
```

Hence we keep the following parameters
```{r vif parameters}
npfsp <- npf.168[,c(1:2,5:6,9:10,13:16,19:20,23:24,27:30,37)]
colnames(npfsp)
npfsp.fact <- npfsp
npfsp.fact[,19] <- as.factor(npfsp.fact[,19])
```

### PCA
#### PCA for all three versions of data 
```{r PCA}
pcaOrig <- prcomp(npf[,1:100], center = T, scale.=T)
cumsum(pcaOrig$sdev^2)/sum(pcaOrig$sdev^2)
x1 <- as.matrix(npf[,1:100])%*%pcaOrig$rotation[,1:12]
pca.168 <- prcomp(npf.168[,1:36], center = T, scale. = T)
cumsum(pca.168$sdev^2)/sum(pca.168$sdev^2)
x2 <- as.matrix(npf.168[,1:36])%*%pca.168$rotation[,1:12]
pca.sp <- prcomp(npfsp[,1:18], center = T, scale. = T)
cumsum(pca.sp$sdev^2)/sum(pca.sp$sdev^2)
x3 <- as.matrix(npfsp[,1:18])%*%pca.sp$rotation[,1:11]
```

@@

## Testing classifiers as they are

@@

```{r naive bayes}
library(e1071)
nb.fit <- naiveBayes(class2 ~ . , data = npf_train, laplace = 1)
nb.pred.cv <- cv.prob(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran

nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2))
# Kaikkiin tulee Inf
nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2))
kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### LDA
```{r lda}
library(MASS)
lda.fit <- lda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
lda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
lda_acc <- data.frame(train = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(lda.pred.cv, npf$class2))
lda_perp <- data.frame(train = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(lda.pred.cv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))%>%
  kable_styling(latex_options = "HOLD_position")
```

### QDA
```{r qda}
library(MASS)
qda.fit <- qda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
qda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
qda_acc <- data.frame(train = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(qda.pred.cv, npf$class2))
# Taas Inf :(
qda_perp <- data.frame(train = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(qda.pred.cv, npf$class2))
kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### GLM
```{r glm, warning=FALSE, fig.height=4}
mod1<-glm(class2~., npf, family=binomial(link="logit"))
mod2<-glm(class2~., npf, family=binomial(link="probit"))
plot(mod1)
prob1<-predict(mod1, npf, type="response")
pred1<-c()
perp1<-c()
for(i in 1:dim(npf)[1]){
  if(prob1[i]>=0.5){
    pred1<-c(pred1,1)
  } else{
    pred1<-c(pred1,0)
  }
  if(npf$class2[i]==1){
    perp1<-c(perp1,prob1[i])
  } else{
    perp1<-c(perp1,1-prob1[i])
  }
}
```

GLM accuracy and perplexity. 
```{r glm res, warning=FALSE}
mean(pred1==npf$class2)
exp(-mean(perp1))
```

### Lasso / Ridge
```{r lasso, message=FALSE, fig.height=4}
library(glmnet)
mod3<-cv.glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=1, family="binomial", type.measure="class")
mod4<-cv.glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=0, family="binomial", type.measure = "class")
mod3a<-glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=1, family="binomial", lambda=mod3$lambda.min)
coef(mod3)
plot(mod3a)
plot(mod3$glmnet.fit, xvar="lambda")
prob3a<-predict(mod3a, newx=as.matrix(npf[,1:100]), type="response")
pred3a<-c()
perp3a<-c()
for(i in 1:dim(npf)[1]){
  if(prob3a[i]>=0.5){
    pred3a<-c(pred3a,1)
  } else{
    pred3a<-c(pred3a,0)
  }
  if(npf$class2[i]==1){
    perp3a<-c(perp3a,prob3a[i])
  } else{
    perp3a<-c(perp3a,1-prob3a[i])
  }
}
```

Lasso accuracy and perplexity. 
```{r lasso  res, warning=FALSE}
mean(pred3a==npf$class2)
exp(-mean(perp3a))
```

```{r ridge, message=FALSE, fig.height=4}
mod4a<-glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=0, family="binomial", lambda=mod4$lambda.min)
plot(mod4a)
plot(mod4$glmnet.fit, xvar="lambda")
prob4a<-predict(mod4a, newx=as.matrix(npf[,1:100]), type="response")
pred4a<-c()
perp4a<-c()
for(i in 1:dim(npf)[1]){
  if(prob4a[i]>=0.5){
    pred4a<-c(pred4a,1)
  } else{
    pred4a<-c(pred4a,0)
  }
  if(npf$class2[i]==1){
    perp4a<-c(perp4a,prob4a[i])
  } else{
    perp4a<-c(perp4a,1-prob4a[i])
  }
}
```

Ridge accuracy and perplexity. 
```{r ridge res, warning=FALSE}
mean(pred4a==npf$class2)
exp(-mean(perp4a))
```

### SVM
```{r svm, warning=FALSE}
modsv<-tune(svm, class2~., data=npf.fact, kernel="linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1, 2, 5)))
modsvm<-svm(class2~., data=npf.fact, kernel="linear", scale=FALSE, cost=0.001,probability=TRUE)
predsv<-predict(modsvm,newdata=npf.fact)
predsv2<-predict(modsvm,newdata=npf.fact,probability = T)
probsv<-attr(predsv2, "probabilities")
perpsv<-c()
for(i in 1:dim(npf)[1]){
  if(npf$class2[i]==0){
    perpsv<-c(perpsv, probsv[i,2])
  } else{
    perpsv<-c(perpsv, probsv[i,1])
  }
}
```
SVM accuracy and perplexity. 
```{r SVM res, warning=FALSE}
mean(predsv==npf$class2)
exp(-mean(log(perpsv)))
```

>>>>

## Accuracy
accuracy <- function(pred, y) mean(ifelse(pred >= 0.5, 1, 0) == y)

## Perplexity
## Force probabilities to [0.995, 0.005]
perplexity <- function(pred, y) {
  pred <- ifelse(pred != 0 & pred != 1, pred, ifelse(pred == 0, 0.005, 0.995 ))
  exp(-mean(log(ifelse(y == 1, pred, 1 - pred)))) 
}

# Cross-validate
# Split n items into k folds of roughly equal size
kpart <- function(n, k) {
    rep_len(1:k, length.out = n)
}

# Cross-validation predictor (from week 1 problem 2)
cv <- function(
    formula, # variables to use
    data, # data
    model = lm, # model to train
    n = nrow(data), # rows in the matrix
    k = min(n, 10), # cross-validation folds
    split = kpart(n, k), # the split of n data items into k folds
    train = function(data) model(formula, data = data),
    pred = function(model, data) predict(model, newdata = data)) {
  yhat <- NULL
  for (i in 1:k) {
    mod <- train(data[split != i, ])
    if (is.null(yhat)) {
      yhat <- pred(mod, data)
    } else {
      yhat[split == i] <- pred(mod, data[split == i, ])
    }
  }
  return(yhat)
}

@@

```{r correlation at different heights, message=FALSE, warning=FALSE}
# Correlation matrix
cm <- cor(npf[, endsWith(colnames(npf), ".mean")])
# Remove .mean from column and row names.
colnames(cm) <- rownames(cm) <- sapply(colnames(cm), 
                                       function(s) gsub(".mean", "", s))
# Correlations for each variables (when more than one variable with the same beginning)
kable(cm[startsWith(colnames(cm), "CO2"), startsWith(colnames(cm), "CO2")],
      caption = "Correlation (CO2)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "H2O"), startsWith(colnames(cm), "H2O")],
      caption = "Correlation (H20)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NO)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NOx)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "O3"), startsWith(colnames(cm), "O3")],
      caption = "Correlation (O3)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "RHIRGA"), startsWith(colnames(cm), "RHIRGA")],
      caption = "Correlation (RHIRGA)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "T"), startsWith(colnames(cm), "T")],
      caption = "Correlation (T)") %>%
  kable_styling(latex_options = "HOLD_position")
```

@@

```{r measurementHeights, message=FALSE, warning=FALSE}
# Selecting which measurements have been done at particular heights
# Columns
cols <- unname(sapply(colnames(npf[endsWith(colnames(npf), ".mean")]), 
                      function(s) gsub(".mean", "", s)))
# Table of variables at different heights (only the variables that have been measured at different heights)
kable(data.frame(dm.42 = 
                   c(sort(cols[endsWith(cols, "42")])[1:6], NA, 
                     sort(cols[endsWith(cols, "42")])[7]), 
                 dm.84 = c(NA, sort(cols[endsWith(cols, "84")])[1:5], NA, 
                           sort(cols[endsWith(cols, "84")])[6]),
                 dm.168 = c(sort(cols[endsWith(cols, "168")])),
                 dm.336 = c(sort(cols[endsWith(cols, "336")])[1:4], NA, 
                            sort(cols[endsWith(cols, "336")])[5], NA, NA),
                 dm.504 = c(sort(cols[endsWith(cols, "504")])[1:6], NA, 
                            sort(cols[endsWith(cols, "504")])[7]),
                 dm.672 = c(NA, NA, sort(cols[endsWith(cols, "672")])[2:5], NA,
                            sort(cols[endsWith(cols, "672")])[6])),
      caption = "Measurements at different heights: What measurements have been done at particular heights?") %>% 
  kable_styling(latex_options = "HOLD_position")
```

### vif
```{r vif, warning=FALSE}
library(car)

testmod <- glm(class2 ~ ., npf.168, family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std, npf.168, 
               family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std -UV_A.std -UV_A.mean, npf.168, 
               family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std -UV_A.std -UV_A.mean -Glob.std 
               -Glob.mean, npf.168, family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std -UV_A.std -UV_A.mean -Glob.std 
               -Glob.mean -UV_B.std -UV_B.mean, npf.168, 
               family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std -UV_A.std -UV_A.mean -Glob.std
               -Glob.mean -UV_B.std -UV_B.mean -RPAR.mean -RPAR.std, npf.168, 
               family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std -UV_A.std -UV_A.mean -Glob.std 
               -Glob.mean -UV_B.std -UV_B.mean -RPAR.mean -RPAR.std -NET.mean 
               -NET.std,  npf.168, family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std -UV_A.std -UV_A.mean -Glob.std 
               -Glob.mean -UV_B.std -UV_B.mean -RPAR.mean -RPAR.std -NET.mean 
               -NET.std -RGlob.std -RGlob.mean, npf.168, 
               family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2 ~ . -PAR.mean -PAR.std -UV_A.std -UV_A.mean 
               -Glob.std -Glob.mean -UV_B.std -UV_B.mean -RPAR.mean -RPAR.std 
               -NET.mean -NET.std -RGlob.std -RGlob.mean -CS.mean -CS.std, 
               npf.168, family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))

testmod <- glm(class2~. -PAR.mean -PAR.std -UV_A.std -UV_A.mean -Glob.std 
               -Glob.mean -UV_B.std -UV_B.mean -RPAR.mean -RPAR.std -NET.mean 
               -NET.std -RGlob.std -RGlob.mean -CS.mean -CS.std -NOx168.mean 
               -NOx168.std, npf.168,family = binomial(link = logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact == max(inflfact))
```

Runnig vif() over several variables in the 16.8m restricted dataset, we come up to the following collection of coefficients 
```{r vif coefficients, warning=FALSE}
coefficients(testmod)
```

Hence we keep the following parameters
```{r vif parameters, warning=FALSE}
npfsp <- npf.168[, c(1:2, 5:6, 9:10, 13:16, 19:20, 23:24, 27:30, 37)]
colnames(npfsp)
npfsp.fact <- npfsp
npfsp.fact[, 19] <- as.factor(npfsp.fact[, 19])
```

### PCA
#### PCA for all three versions of data 
```{r PCA, warning=FALSE}
pcaOrig <- prcomp(npf[, 1:100], center = T, scale. = T)
cumsum(pcaOrig$sdev^2)/sum(pcaOrig$sdev^2)
x1 <- as.matrix(npf[, 1:100]) %*% pcaOrig$rotation[, 1:12]
pca.168 <- prcomp(npf.168[, 1:36], center = T, scale. = T)
cumsum(pca.168$sdev^2)/sum(pca.168$sdev^2)
x2 <- as.matrix(npf.168[, 1:36]) %*% pca.168$rotation[, 1:12]
pca.sp <- prcomp(npfsp[, 1:18], center = T, scale. = T)
cumsum(pca.sp$sdev^2)/sum(pca.sp$sdev^2)
x3 <- as.matrix(npfsp[, 1:18]) %*% pca.sp$rotation[, 1:11]

npf.pca <- as.data.frame(cbind(x1, npf$class2))
colnames(npf.pca) <- c(colnames(x1), "class2")
npf.168.pca <- as.data.frame(cbind(x2, npf.168$class2))
colnames(npf.168.pca) <- c(colnames(x2), "class2")
```

@@

## Naive Bayes, LDA and QDA on the original data, 168-data and the PCA'd versions of these data sets

@@

```{r naive bayes, warning=FALSE}
library(e1071)

# Original data
nb.fit <- naiveBayes(class2 ~ . , data = npf_train)
nb.pred.cv <- cv.prob(class2 ~ ., npf, naiveBayes)       # Very slow so done only once
nb.pred.loocv <- loocv.prob(class2 ~ ., npf, naiveBayes) # Same as above

nb_acc <- data.frame(
  train = accuracy(predict(nb.fit, newdata = npf_train, 
                           type = "raw")[, 2], npf_train$class2),
  test = accuracy(predict(nb.fit, newdata = npf_test, 
                          type = "raw")[, 2], npf_test$class2),
  CV = accuracy(nb.pred.cv, npf$class2),
  LOOCV = accuracy(nb.pred.loocv, npf$class2))

nb_perp <- data.frame(
  train = perplexity(predict(nb.fit, newdata = npf_train, 
                             type = "raw")[, 2], npf_train$class2),
  test = perplexity(predict(nb.fit, newdata = npf_test, 
                            type = "raw")[, 2], npf_test$class2),
  CV = perplexity(nb.pred.cv, npf$class2),
  LOOCV = perplexity(nb.pred.loocv, npf$class2))
kable(cbind(nb_acc, nb_perp)) %>%
  add_header_above(c("Accuracy" = 4, "Perplexity" = 4)) %>%
  kable_styling(latex_options = "HOLD_position")

# 168-data
nb.pred.cv.168 <- cv.prob(class2 ~ ., npf.168, naiveBayes)       # Very slow so done only once
nb.pred.loocv.168 <- loocv.prob(class2 ~ ., npf.168, naiveBayes) # Same as above

nb_acc.168 <- data.frame(CV = accuracy(nb.pred.cv.168, npf.168$class2),
                        LOOCV = accuracy(nb.pred.loocv.168, npf.168$class2))

nb_perp.168 <- data.frame(CV = perplexity(nb.pred.cv.168, npf.168$class2),
                          LOOCV = perplexity(nb.pred.loocv.168, npf.168$class2))
kable(cbind(nb_acc.168, nb_perp.168)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

Using PCA on the original and the 16.8m dataset.

```{r naive bayes pca, warning=FALSE}

# Original data
nb.pred.cv.pca <- cv.prob(class2 ~ ., npf.pca, naiveBayes)       # Very slow so done only once
nb.pred.loocv.pca <- loocv.prob(class2 ~ ., npf.pca, naiveBayes) # Same as above

nb_acc.pca <- data.frame(CV = accuracy(nb.pred.cv.pca, npf.pca$class2),
                         LOOCV = accuracy(nb.pred.loocv.pca, npf.pca$class2))

nb_perp.pca <- data.frame(CV = perplexity(nb.pred.cv.pca, npf.pca$class2),
                          LOOCV = perplexity(nb.pred.loocv.pca, npf.pca$class2))
kable(cbind(nb_acc.pca, nb_perp.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

# 168-data
nb.pred.cv.168.pca <- cv.prob(class2 ~ ., npf.168.pca, naiveBayes)       # Very slow so done only once
nb.pred.loocv.168.pca <- loocv.prob(class2 ~ ., npf.168.pca, naiveBayes) # Same as above

nb_acc.168.pca <- data.frame(CV = accuracy(nb.pred.cv.168, npf.168.pca$class2),
                     LOOCV = accuracy(nb.pred.loocv.168, npf.168.pca$class2))

nb_perp.168.pca <- data.frame(CV = perplexity(nb.pred.cv.168, npf.168.pca$class2),
                      LOOCV = perplexity(nb.pred.loocv.168, npf.168.pca$class2))
kable(cbind(nb_acc.168.pca, nb_perp.168.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

### LDA

On the original dataset and the 16.8m dataset.
```{r lda, warning=FALSE}
library(MASS)

# Cross-validation parameters
n = nrow(npf)       # number of rows in the data matrix
k = min(n, 10)      # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds

lda.pred.cv <- rep(0,n)
lda.pred.cv.168 <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
  ## 168
  mod <- lda(class2 ~ ., npf.168[split != i, ])
  lda.pred.cv.168[split == i] <- predict(mod, npf.168[split == i, ])$posterior[,2]
}

lda.pred.loocv <- rep(0, n)
lda.pred.loocv.168 <- rep(0, n)

# loocv
for (i in 1:dim(npf)[1]) {
  mod <- lda(class2 ~ ., npf[-i, ])
  lda.pred.loocv[i] <- predict(mod, npf[i, ])$posterior[,2]
  ## 168
  mod <- lda(class2 ~ ., npf.168[-i, ])
  lda.pred.loocv.168[i] <- predict(mod, npf.168[i, ])$posterior[,2]
}

# npf
lda_acc <- data.frame(CV = accuracy(lda.pred.cv, npf$class2),
                      LOOCV = accuracy(lda.pred.loocv, npf$class2))
lda_perp <- data.frame(CV = perplexity(lda.pred.cv, npf$class2),
                       LOOCV = perplexity(lda.pred.loocv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

# 168
lda_acc.168 <- data.frame(CV = accuracy(lda.pred.cv.168, npf$class2),
                          LOOCV = accuracy(lda.pred.loocv.168, npf$class2))
lda_perp.168 <- data.frame(CV = perplexity(lda.pred.cv.168, npf$class2),
                           LOOCV = perplexity(lda.pred.loocv.168, npf$class2))
kable(cbind(lda_acc.168, lda_perp.168)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")
```

Using PCA on the original and the 16.8m dataset.

```{r lda pca, warning=FALSE}

# Cross-validation parametrit
n = nrow(npf)       # number of rows in the data matrix
k = min(n, 10)      # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds

lda.pred.cv.pca <- rep(0,n)
lda.pred.cv.168.pca <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf.pca[split != i, ])
  lda.pred.cv.pca[split == i] <- predict(mod, npf.pca[split == i, ])$posterior[,2]
  ## 168
  mod <- lda(class2 ~ ., npf.168.pca[split != i, ])
  lda.pred.cv.168.pca[split == i] <- predict(mod, npf.168.pca[split == i, ])$posterior[,2]
}

lda.pred.loocv.pca <- rep(0, n)
lda.pred.loocv.168.pca <- rep(0, n)

# loocv
for (i in 1:dim(npf)[1]) {
  mod <- lda(class2 ~ ., npf.pca[-i, ])
  lda.pred.loocv.pca[i] <- predict(mod, npf.pca[i, ])$posterior[, 2]
  ## 168
  mod <- lda(class2 ~ ., npf.168.pca[-i, ])
  lda.pred.loocv.168.pca[i] <- predict(mod, npf.168.pca[i, ])$posterior[, 2]
}

# npf
lda_acc.pca <- data.frame(CV = accuracy(lda.pred.cv.pca, npf.pca$class2),
                          LOOCV = accuracy(lda.pred.loocv.pca, npf.pca$class2))
lda_perp.pca <- data.frame(CV = perplexity(lda.pred.cv.pca, npf.pca$class2),
                           LOOCV = perplexity(lda.pred.loocv.pca, npf.pca$class2))
kable(cbind(lda_acc.pca, lda_perp.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

# 168
lda_acc.168.pca <- data.frame(CV = accuracy(lda.pred.cv.168.pca, 
                                            npf.168.pca$class2),
                              LOOCV = accuracy(lda.pred.loocv.168.pca, 
                                               npf.168.pca$class2))
lda_perp.168.pca <- data.frame(CV = perplexity(lda.pred.cv.168.pca, 
                                               npf.168.pca$class2),
                               LOOCV = perplexity(lda.pred.loocv.168.pca, 
                                                  npf.168.pca$class2))
kable(cbind(lda_acc.168.pca, lda_perp.168.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2) )%>%
  kable_styling(latex_options = "HOLD_position")
```

### QDA
```{r qda, warning=FALSE}
library(MASS)

# Cross-validation parametrit
n = nrow(npf)       # number of rows in the data matrix
k = min(n, 10)      # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds

qda.pred.cv <- rep(0, n)
qda.pred.cv.168 <- rep(0, n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[, 2]
  ## 168
  mod <- qda(class2 ~ ., npf.168[split != i, ])
  qda.pred.cv.168[split == i] <- predict(mod, npf.168[split == i, ])$posterior[, 2]
}

qda.pred.loocv <- rep(0, n)
qda.pred.loocv.168 <- rep(0, n)
